manuscript_removed_text

### 2019-09-12 manuscript/manuscript_removed_text.Rmd
Update manuscript_removed_text.Rmd

>>>>

In this analysis we used a comprehensive global database, containing an unprecedented amount of data and representing all global forest ecosystems and all significant forest carbon fluxes, to comprehensively explore the relationships between climate and productivity on a global scale. Many of our findings support and clarify previously published work; however we used a much larger and more complete database than has previously been available, allowing for higher quality control and standardisation of data. This enabled us to control for factors not previously controlled for - such as stand age, methodology, flux components, and disturbance regime - to gain a stronger understanding of latitudinal and climatic effects on productivity.

Our results show that productivity decreases linearly with latitude. Climate explains a significant proportion of variation in all carbon fluxes, with temperature variables being the best predictors of productivity on this global scale. While other climate variables are significant predictors of productivity, none of them improve on the explanatory power of temperature-related variables in general or MAT specifically. Water availability is an important factor in explaining productivity on a global scale: we find a positive influence of precipitation at low MAP, with saturation at higher levels of MAP. There is a significant interaction between MAT and MAP.  We note that climate tends to explain a higher amount of variation in the major fluxes, with the amount of variation explained decreasing in smaller component fluxes.

We find that seasonality is an important factor in understanding patterns in productivity on a global scale: temperature seasonality and growing season length are strong predictors of productivity, though growing season length doesn't improve on MAT as a predictor. Within the growing season we find the influence of climate on productivity is smaller, but still significant for a number of carbon fluxes.

Despite the high standard of data quality in the ForC database, there are still significant limitations on our analyses. We standardised methodologies as far as possible, ensuring that there was consistency in the components included for each carbon flux. However, field techniques for estimating components of primary productivity are variable in their accuracy. This is particularly true for measures of BNPP, which are often estimated by extrapolating measures of ANPP. Even where BNPP is calculated through direct soil sampling, measures are often taken to insufficient depth, leading to underestimation of the flux. It is challenging to account for this type of variability in quality of field data, and our analyses largely assume good faith in data quality. While we have strong confidence in our analyses of well-understood fluxes, such as GPP, ANPP, and ANPP~woody~ ~stem~, we recommend caution in drawing firm conclusions from our analyses of BNPP, BNPP~fine~ ~root~, and, to some extent, NPP. We expect that, as field techniques improve the accuracy of estimates of these fluxes in the future, our understanding of the relationship of BNPP with climate variables will change. It is possible that the low R^2^ values for BNPP that we find are a result of lower accuracy in field measures of these fluxes, and that with improve standardisation of methods, we would find climate explains a higher proportion of variation.

Secondly, we emphasise that this study considers the relationship between productivity and climate on a coarse scale. Because of the nature of the ForC database, we were unable to access site-level MAT and MAP data for a number of sites, and lacked site-level data for other climate variables for all sites. As a result the climate data we used was primarily multi-year averages of annual climate measures accessed from global databases. While the quality of these data is high, we cannot expect that it captures detailed local variation in climate, and in particular will be unable to capture microclimate variation, which may have a strong influence on stand-level productivity. Although we find that MAT is the strongest predictor for climate, part of this effect may be a result of the higher-quality stand-level dataset available for MAT. With high-quality stand-level data for other climate variables we expect that their predictive strength would improve.

Furthermore, we expect that temporal resolution of climate data affects our results. Productivity is known to vary on a seasonal and annual scale in response to variation in climate. Our analysis relies on long-term averages in both climate variables and forest productivity estimates, and as a result averages out a large proportion of this variation. Therefore we recognise that the strong short-term effects of climate on productivity are likely to be masked. The limitations of low temporal resolution are highlighted in our analysis of growing season length. We note that our estimate of growing season length is imprecise, being based on monthly temperature, precipitation, and evapotranspiration averages. With accurate stand-level measurements of growing season length, based on daily, not monthly, climate measures, we would expect to record a stronger correlation between growing season length and productivity.

Although we recognise the inherent limitations present in this analysis, by focusing on mature, undisturbed forests, and by clarifying the components of NPP included in analysis, our study offers a clearer picture of how carbon cycling varies with latitude and climate.

Past studies have differed in their conclusions regarding the relationship between productivity and latitude. Our findings indicate that, for all carbon fluxes studied, productivity peaks in the tropical regions. There is strong evidence from our analyses that this increase in productivity towards the tropical regions is driven primarily by temperature-related effects. We find that temperature-related climate variables explain the highest proportion of variability in productivity. In contrast, we find that MAP is not a significant predictor of productivity, but does show a strong interaction with temperature. This indicates that the main mechanism through with precipitation influences productivity is in a secondary capacity, with its overall influence being mediated by the effects of temperature. This also suggests that although temperature explains a large proportion of variation, it cannot fully explain productivity alone, and a measure of water availability is necessary to fully understand patterns of productivity. This is likely to be significant in the future, as it suggests that the effects of a warming climate will be mediated through changes in precipitation regime.

There is debate over whether productivity in regions with higher MAT is higher because productivity inherently increases with air temperature during growing season months, or because higher MAT is correlated with longer growing seasons. Our analysis suggests that length of growing season is a key factor in understanding latitudinal patterns in productivity. We find that length of growing season is strongly correlated with productivity. This is supported by the strong correlation between temperature seasonality and productivity, indicating that aseasonality and year-round growing seasons have a strong influence over productivity.

These conclusions are supported by our analyses of the influence of climate within the growing season. We find much weaker correlations between climate variables and productivity when only growing season months are considered. There is evidence for a positive correlation between productivity and temperature for ANPP and ANPP~foliage~, suggesting that there is still a small positive effect of air temperature on productivity independent of growing season, however this effect is not significant for the remaining fluxes. This suggests that the primary positive effect of temperature on productivity is through the lengthening of the growing season in regions with higher MAT. However, we note that length of growing season is a less good predictor of productivity than MAT. As discussed above, this may be an artefact of imprecise estimation of length of growing season, however it does indicate that there are aspects of productivity explained by MAT which length of growing season alone cannot explain.

By including multiple fluxes in our analysis, we are able to explore the relationships between fluxes and their responses to climate. We find that fluxes are broadly consistent in their responses to climate drivers, suggesting that the importance of climate drivers on productivity is relatively invariant across fluxes. In addition, we show that carbon allocation between subsidiary fluxes does not vary significantly across broad climatic gradients. There is evidence that allocation between carbon fluxes does vary with factors such as stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill_belowground_2016], forest structure [@taylor_greater_2019], elevation [@moser_elevation_2011], and water availability [@newman_above-_2006]. That we do not see consistent patterns of variation in allocation across climatic gradients indicates that the pressures influencing shifts in allocation between fluxes occur on local scales, and that shifts in allocation occur as a response to fluctuations in local environmental conditions.

Related to this, we find that climate more strongly explains variation in the major fluxes, with lower levels of variation explained in subsidiary fluxes. Although this could in part be related to a lack of standardisation in methodology across studies, as discussed above, it could also indicate that factors other than climate have significant influence over subsidiary fluxes. This is consistent with the results we find for allocation between fluxes: although climate has a significant influence on subsidiary fluxes, it is not the only factor that is important, and many other local- and regional-scale factors  influence the way carbon is allocated between subsidiary fluxes. We would expect that this would reduce the strength of the effect of climate on subsidiary fluxes.

**[needs strong concluding paragraph]**
